### 2019-06-24 literature-review/explaining-mpe.tex
inizio sezione 3.3

% !TEX root = thesis-thomas-tiotto.tex

\section{Explaining the Most Probable Explanation}\label{sec:explaining-mpe}
\cite{Butz2018}