### 2019-07-08 mathematical-background/graph-theory.tex

\section{Graph Theory} \label{sec:graph-theory}
Many problems in Machine Learning (ML) don't involve classification or prediction of single data points in isolation, but of set of entities that may present a more, or less, complex relation with each other. 
Most real-world phenomena fit into the latter framework.
Graphs are one of the most powerful tools for the modelling of this class of problems, as their structure naturally captures the wide variety of relations that may exist between entities.
These range from the atomical structure of a molecule to a social network of friends.  
In both these examples graphs help in reasoning, visualising and making inferences and predictions.

\subsection{Graphs} \label{subsec:graphs}
\begin{definition}
	A graph is a tuple 
	\begin{equation}
	\mathcal{G} = (\mathcal{V}, \mathcal{E})
\end{equation}
with $\mathcal{V} = \{ v_1 \ldots v_n \}$ the set of \textit{vertices} and $\mathcal{E} = \mathcal{V} \times \mathcal{V}$ the set of \textit{edges}.
\end{definition}

For our scopes, we will only be considering the case where every element in $\mathcal{E}$ is a pair either of the form $(v_i, v_j)$ or $(v_j, v_i)$ with $i \neq j$.  
That is to say that the class of graphs presently of interest for us are those where there can be at most a single directed edge between any node in $\mathcal{V}$ and no self-loops.
We are also interested in enforcing that there be no \textit{cycles} in the graph, i.e. sequences of nodes of the form $v_i \rightarrow v_j \rightarrow \cdots \rightarrow v_i$.
The resulting graph possessing only directed edges and no cycles is commonly called a \textit{directed acyclic graph}, or DAG for short.  
This data structure is of paramount importance as it's the fundamental graphical representation used for Bayesian Networks.

\subsection{Polytrees} \label{subsec:polytrees}
We now have all elements to be able to formally define a Bayesian Network.
I will also define polytrees and trees because these are a fundamental concept for the work carried out in this thesis.
\begin{definition}
	A \textit{loop} is a trace $v_i, v_j \ldots v_i$ of nodes obtained by following edges regardless of their direction
\end{definition}
\begin{definition}
	A directed graph containing no such loops is called a \textit{polytree}. 
\end{definition}
\begin{definition}
	A \textit{tree} is a particular case of polytree where each node has at most one parent.	
\end{definition}

\subsection{D-separation} \label{subsec:d-separation}
\textit{Dependence-separation} or \textit{d-separation}, as the name entails, is a concept relating to the conditional dependence between variables.
It was first presented by \cite{Pearl1988}
To define it, we first have to clarify when two sets of nodes $X$ and $Y$ are causally connected.
This is so if $Z = \emptyset$ and they are part of one of the following three structures, called \textit{v-structures} in this context:
\begin{itemize}
  \item $X \rightarrow Z \rightarrow Y$
  \item $X \leftarrow Z \leftarrow Y$
  \item $X \leftarrow Z \rightarrow Y$
\end{itemize}
This means that knowing something about $X$ also tells us something new about $Y$.
$X$ and $Y$ are causally independent if they appear in the following v-structure:
\begin{itemize}
  \item $X \rightarrow Z \leftarrow  Y$
\end{itemize}
Such a configuration is called a \textit{collider} and it blocks the flow of information from $X$ to $Y$.
If $Z \neq \emptyset$ the cases are reversed so colliders are open and the other three structures are blocked.
\begin{definition}
	Given disjoint subsets $X, Y, Z \subset \mathcal{X}$, $X$ and $Y$ are d-separated if:
	\begin{itemize}
		\item $Z \neq \emptyset$: no path between $X$ and $Y$ presents a collider
		\item $Z = \emptyset$: there is a collider on every path between $X$ and $Y$
	\end{itemize}
\end{definition}

The independencies between variables are encoded in the structure of the DAG so every distribution whose BN has the same connections between nodes also has the same independencies, regardless of the values of the variables.

An example using on the real network used in this thesis is shown in Fig. \ref{fig:d-separation_example}.
Here we can see how the network's topology and the nodes chosen to be in the observed set $Z$ define the resulting separations.
In this case $X= \{ \text{mut17q21} \}$, $Z= \{ \text{differenziazione, pN} \}$ and $Y=V \smallsetminus X \smallsetminus Z$; the resulting set of nodes that are d-separated from $X$ contains $\{ \text{c erB 2, FISH, loss 17, situ, pT, M, lateralita} \}$.
Basically, we are asking for the set of all nodes in the DAG that are d-separated from $X$, given evidence $Z$.
The reason for this can easily be given by enumerating all paths through the v-structures in the network and applying the definitions for causal connections given above:
\begin{itemize}
  \item mut17q21 $\rightarrow$ eta arrotondata: are directly connected so the cannot be independent
  \item mut17q21 $\rightarrow$ recettori estrogeni: like above
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ morfologia: recettori estrogeni is not in $Z$ so the path is open and mut17q21 is d-connected to morfologia
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ recettori progestinici: like above
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ ki67: as above
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ differenziazione: as above
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ differenziazione $\rightarrow$ c erbB 2: differenziazione is in $Z$ so this path is closed, as there are no other paths between mut17q21 and c erbB 2, the two nodes are d-separated
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ differenziazione $\rightarrow$ FISH: the only path to mut17q21 is through c erbB 2, that is already d-separated from mut17q21, so the two nodes are d-separated
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ differenziazione $\rightarrow$ FISH $\rightarrow$ loss 17: as above
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ differenziazione $\rightarrow$ pN: as recettori estrogeni, differenziazione and pN are arranged in a \textit{collider} and differenziazione is in $Z$, then the path is open
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ differenziazione $\rightarrow$ pN $\rightarrow$ situ: pN is in $Z$ and differenziazione, pN, situ are arranged as a \textit{fork}, thus the path is closed
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ differenziazione $\rightarrow$ pN $\rightarrow$ pT: as above, M is also d-separated from mut17q21 so there are no open paths between mut17q21 and pN
  \item mut17q21 $\rightarrow$ recettori estrogeni $\rightarrow$ differenziazione $\rightarrow$ pN $\rightarrow$ M: differenziazione, pN, M are arranged as a \textit{chain} and pN is in $Z$, the path is closed
  \item lateralita: lateralita is not connected to any other node and is thus d-separated from all others a priori
\end{itemize}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{methodology/images/d-separation-example}}
\caption{D-Separations in the provided data set (see Sec. \ref{sec:data-set}) generated using the \texttt{plot\_model} function (see Subsec. \ref{subsubsec:plot-model})}
\label{fig:d-separation_example}
\end{figure}