### 2019-07-18 results/implemented-tool.tex

\section{Implemented tool}
\subsection{Overview}

\subsection{Independencies query}

\subsection{Conditional probability query}

\subsection{MPE query}

\subsection{Pseudo-MPE query}

\subsection{Exhaustive dialogue}

\subsection{Independencies dialogue}

\subsection{Threshold dialogue}

\subsection{Pseudo-MPE comparison}