### 2019-08-28 thesis-thomas-tiotto.tex

\documentclass[mscthesis]{usiinfthesis}
\usepackage{lipsum}
\input{macros}

%\usepackage[style=numeric,natbib=true]{biblatex}

\usepackage{tikz}
\usetikzlibrary{graphs}

\usepackage{natbib}

\usepackage{listings}
\newcommand{\Eta}{H}

\usepackage{tabularx}
\usepackage{ragged2e}
\newcolumntype{Y}{>{\RaggedRight\arraybackslash}X} 
\usepackage{booktabs}
\renewcommand\tabularxcolumn[1]{m{#1}}

\usepackage{algorithm, algorithmicx, algpseudocode}

\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{definition}[theorem]{Definition} 
\newtheorem{corollary}[theorem]{Corollary}

\lstdefinelanguage{algebra}
{morekeywords={import,sort,constructors,observers,transformers,axioms,if,
else,end},
sensitive=false,
morecomment=[l]{//s},
}

\usepackage{csquotes}

\usepackage{fge}
\newcommand{\mysetminus}{\mathbin{\fgebackslash}}

\usepackage{rotating}
\usepackage{multirow}
\usepackage{caption}

\usepackage{nameref}

\usepackage{framed}
\usepackage{enumitem}

\usepackage{centernot}

\usepackage{pgfplots}

\usepackage{cancel}
\usepackage{soul}

\usepackage[section]{placeins}

\title{Explainable AI with Probabilistic Graphical Models} %compulsory 
\specialization{Artificial Intelligence}%optional
\subtitle{A Study with Application and Evaluation on a Medical Domain} %optional 
\author{Thomas Francesco Tiotto} %compulsory
\begin{committee}
\advisor{Dr.}{Alessandro Facchini}{} %compulsory
\coadvisor{Dr.}{Alessandro Antonucci}{}{} %optional
\end{committee}
\Day{6} %compulsory
\Month{September} %compulsory
\Year{2019} %compulsory, put only the year
\place{Lugano} %compulsory

\dedication{To everyone} %optional
\openepigraph{Considerate la vostra semenza:\\
fatti non foste a viver come bruti,\\
ma per seguir virtute e canoscenza.}{Dante Alighieri} %optional

%\makeindex %optional, also comment out \theindex at the end

\begin{document}

\maketitle %generates the titlepage, this is FIXED

\frontmatter %generates the frontmatter, this is FIXED

\begin{abstract}
\input{abstract}
\end{abstract}

\begin{acknowledgements}
First and foremost, I would like to thank my supervisors Dr.\ Alessandro Facchini and Dr.\ Alessandro Antonucci for being patient and enthusiastic enough to follow me through all this process.
There is no way I could have done all this without them and I thoroughly enjoyed all my the interactions with them.

Likewise, Dr.\ Vittoria Martin from Istituto Cantonale di Patologia was invaluable in providing her honest support and often went beyond what could be expected.
The same can be said of Dr.\ Ginevra Licandro, whose insights were essential in helping me to bridge the gap between the Computer Science and Medical domains.

I would naturally want to thank my family and girlfriend for all the support they have always given me and understanding they have had, especially in this final period of my studies.
Their fortitude and unwavering presence has helped in ferrying me across all rough waters.
\end{acknowledgements}

\tableofcontents 
\listoffigures %optional
\listoftables %optional

\mainmatter

%%%%
%%%% INTRODUCTION
%%%%
\chapter{Introduction}\label{chap:introduction}
\input{introduction/context}
\input{introduction/problem}
\input{introduction/response}

%%%%
%%%% LITERATURE REVIEW
%%%%  
\chapter{Literature Review}\label{chap:literature-review}
\input{literature-review/introduction}
\input{literature-review/explainability}
\input{literature-review/explaining-mpe}
\input{literature-review/summary}

%%%%
%%%% MATHEMATICAL BACKGROUND
%%%%
\chapter{Mathematical Background}\label{chap:mathematical-background}
\input{mathematical-background/introduction}
\input{mathematical-background/probability-theory}
\input{mathematical-background/information-theory}
\input{mathematical-background/graph-theory}
\input{mathematical-background/bayesian-networks}
\input{mathematical-background/summary}

%%%%
%%%% METHODOLOGY
%%%%
\chapter{Methodology} \label{chap:methodology}
\input{methodology/introduction}
\input{methodology/data-set}
\input{methodology/instruments}
\input{methodology/novel-contributions}
\input{methodology/validation}
\input{methodology/summary}

%%%%
%%%% RESULTS
%%%%
\chapter{Results} \label{chap:results}
\input{results/introduction}
\input{results/implemented-tool}
\input{results/validation-results}
\input{results/pseudo-mpe-evaluation}
\input{results/prediction-evaluation}
\input{results/issues}
\input{results/summary}

%%%%
%%%% CONCLUSIONS
%%%%
\chapter{Conclusions} \label{chap:conclusions}
\input{conclusions/discussion}
\input{conclusions/future-work}

%%%%
%%%% CONCLUSIONS
%%%%
\appendix %optional, use only if you have an appendix

\backmatter

\chapter{Acronyms} \label{chap:acronyms} %optional
\input{acronyms}

\chapter{Annex} \label{chap:annex}
\input{annex}

%\bibliographystyle{alpha}
%\bibliographystyle{dcu}
\bibliographystyle{plainnat}
\bibliography{biblio}

%\cleardoublepage
%\theindex %optional, use only if you have an index, must use
	  %\makeindex in the preamble

\end{document}